Individual heterogeneity in functional topography of the default-mode network in medial prefrontal cortex

### 2019-04-04 Manuscript.Rmd
working on TBDs (switching to medians, etc)

# In case I want to visualize stuff in the brain
labelCoords_vertex <- read.csv2('labelCoords_vertex.csv')[,2:6]
labelCoords_vertex <- transform(labelCoords_vertex, x = as.numeric(as.character(x)), y = as.numeric(as.character(y)), z = as.numeric(as.character(z)))

@@

Figure 2 shows a representative partitioning of the search space for a single participant (100307). The SP algorithm subdivides a network according to the positive versus negative values in the Fiedler vector (the eigenvector related to the second-to-lowest eigenvalue of the network's Laplacian matrix, see Methods). Since this is a data-driven approach, there is no a priori labeling for the two communities. We thus defined the DMN community as the one containing the one DMN-specific PCC parcel from the meta-analysis (7m). In qualitative terms, the resulting patterns contained substantial DMN coverage in posterior PCC (as dictated by our labeling strategy), with non-DMN vertices in anterior PCC. The mPFC region tended to include DMN vertices in its ventral-anterior and dorsal-anterior areas, with a persistent non-DMN pocket between them. This non-DMN section extended posteriorly into pregenual cinglate cortex (area a24).

@@

# Median size of the extra (>2) communities estimated by modularity
modularityCounts <- lapply(Summaries, function(data) {sort(count(data$Membership)$freq, decreasing = T)})
modularityCounts_two <- sapply(modularityCounts, function(x) {mean(x[seq(2)], na.rm = T)})
modularityCounts_extra <- sapply(modularityCounts, function(x) {mean(x[3:4], na.rm = T)})

@@

This comparison showed that both clustering methods were in high agreement across subjects (mean ARI = `r round(mean(RI_all_within$Index), digits = 2)`, SD = `r round(sd(RI_all_within$Index), digits = 2)`). Modularity showed a tendency to produce additional communities (median = `r median(nModularity)`, max = `r max(nModularity)`). However, these communities encompassed a **(TBD: small number of vertices (mean = `r mean(modularityCounts_extra, na.rm = T)`, IQR = `r IQR(modularityCounts_extra, na.rm = T)`) compared to the principal two (mean = `r mean(modularityCounts_two, na.rm = T)`, IQR = `r IQR(modularityCounts_two, na.rm = T)`)**, suggesting that the overall network is composed of two main communities. In addition, the estimation of the optimal number of cuts based on the rank of the first non-zero eigenvalue (see "Partition evaluation" for details) showed that `r prcntDissection` percent of the individual networks were best partitioned into two communities.

@@

### *Pattern stability over time*

@@

There is precedent in thhe literature for the idea that the magnitude (and not just the sign) of the Fiedler vector values conveys important information about the role of each node in the network [@Gkantsidis2003; @Tian2018]. Therefore, we asked whether the magnitude of these eigenvector values was associated with the stability of nodes in time. Specifically, we estimated the proportion of DMN affiliations per node as a function of Fiedler vector values, using a logistic mixed effects model (Figure 4). The model yielded a positive significant relationship between these features ($\beta$ = `r round(SC_FVvPDMN_modelSummary$coefficients[2,1], digits = 2)`, SE = `r round(SC_FVvPDMN_modelSummary$coefficients[2,2], digits = 2)`, *p* < 0.001), meaning that vertices with **higher absolute Fiedler vector values** were more persistent in their relationship with their corresponding community over time. We refer to this proportion of affiliations over time as the stability of a given node.

@@

Once again we found low alignment across individuals for PCC (mean = `r round(mean(Halves_PCCComparison[[2]]), digits = 2)`, SD = `r round(sd(Halves_PCCComparison[[2]]), digits = 2)`) and mPFC (mean = `r round(mean(Halves_PFCComparison[[2]]), digits = 2)`, SD = `r round(sd(Halves_PFCComparison[[2]]), digits = 2)`), but both areas showed comparatively high levels of within-individual agreement (PCC: mean = `r round(mean(Halves_PCCComparison[[1]]), digits = 2)`, SD = `r round(sd(Halves_PCCComparison[[1]]), digits = 2)`; mPFC: mean = `r round(mean(Halves_PFCComparison[[1]]), digits = 2)`, SD = `r round(sd(Halves_PFCComparison[[1]]), digits = 2)`). We calculated an index of relative specificity by computing the ratio of each individual’s across-day (within-participant) ARI to the mean of all between-participant ARI values involving that individual. **The index is expected to take on a value near 1 if partitionings are well aligned across individuals and/or are subject to a common level of measurement noise. It is expected to exceed 1 insofar as functional network organization is reliable and individual-specific. This index is intended to factor out the potential effects of measurement noise, as they are asumed to be constant for each region (TBD, not sure if needed).**

Figure 6 shows these ARI ratios for PCC and mPFC for all individuals As expected, we found that while both regions showed evidence for specificity (i.e. ratios > 1), the ratios for mPFC (median = `r round(median(Halves_RIRatios_PFC), digits = 2)`; SD = `r round(sd(Halves_RIRatios_PFC), digits = 2)`) were significantly higher than those for PCC (median = `r round(median(Halves_RIRatios_PCC), digits = 2)`; SD = `r round(sd(Halves_RIRatios_PCC), digits = 2)`) when compared with a paired permutation test (*p* `r ifelse(Halves_mPFCvPCC_permTest$Pval < 0.001, "< 0.001", paste("=", Halves_mPFCvPCC_permTest$Pval))`; Cohen's D = `r round(Halves_mPFCvPCC_ES, digits = 2)`) **(TBD: test against 1. I think it might be worth doing it for PCC's sake, as mPFC's numbers + figure already convey its idiosyncrasy very well. If we do it here, we'll do it for the run-specific analysis too)**. These test/retest results suggest that the topographic variability seen in mPFC arises at least in part from stable and subject-specific organizational patterns.

@@

The computational demands of our approach escalate quickly as the number of vertices increases. We thus explored the possible advantage of community detection relative to a simple seed-based functional connectivity analysis for estimating the individual-specific functional topography of mPFC. We examined whether maps generated with SP were more similar per participant across days than those computed from seed-based correlations. We generated a seed time-series by averaging all vertices in the PCC region of our search space, and calculated its correlation with the activity of each vertex in mPFC **(TBD: limiting this to 7m as a seed. I think it makes sense, especially since we're saying that PCC (1) is also subdivided to some extent, and (2) also shifts across individuals to some extent. Alternatively, you wouldn't use 7m as a seed unless you just wanted to find DMN. Full PCC is what a traditional analysis might have done.)**. We compared the map of correlation values in mPFC to the map of unbinarized Fiedler vector values. We performed pairwise spatial Spearman correlations among maps computed for each day and method from all subjects. For `r nrow(unCorrParticipants %>% filter(Corrs < 0))` subjects, the maps produced with one of the days' datasets insufficiently covered area 7m, and so the community labeling was reversed (for one additional subject, both daily maps were reverted). ARI is robust to these labeling issues, but this produced strong negative correlations of the Fiedler vector for these subjects, even though the topography was highly similar. We thus visually inspected these subjects (n = 7) and reverted the labeling when it was clear that the topologies aligned across days. Figure 7 (left) shows that these pairwise comparisons resemble those from the day-to-day comparisons above, and suggests good alignment between methods, but particularly high agreement within subject and method.

@@

The thresholded partitions in our work qualitatively resembled the topographical arrangement of DMN sub-networks A and B proposed by Braga and Buckner [-@Braga2017]. We explored this similarity by reproducing their seed-based connectivity approach in three of our subjects. In their work, Braga and Buckner selected individual vertices along the dorsolateral prefrontal cortex (DLPFC) that produced two anticorrelated networks. Therefore, we hypothesized that if the communities we identified on the medial wall matched their proposed networks, we should be able to find seed vertices in DLPFC that can reproduce them. The whole-brain seed-based functional connectivity maps for these subjects are juxtaposed to their spectral partitioning counterparts on figure 9.

Visual inspection of these networks shows a remarkable similarity between our DMN and non-DMN communities and the previously reported sub-networks A and B. Our search space does not comprise a number of diagnostic areas used in identifying netowrk A, such as the temporo-parietal junction and parahippocampal area [@Braga2019]. Lack of signal from these areas could drive the small discrepancies in the topography between our work and theirs. Nevertheless, these qualitative similarities are important, as they collectively suggest there may exist a true functional partitioning of canonical DMN regions into two distinct communities. Results from the present investigation suggest that the topographic organization of the communities is variable across a large sample of participants, especially in mPFC, and that the SP algorithm is a useful and reliable tool for identifying the two communities in individuals.

@@

### *Table 1. Parcels from Glasser et al. (2016) contained in the search space.*
| Hemisphere  |            mPFC        |          PCC          |
| :---------- | :--------------------: | --------------------: |
| Left        | `r searchSpace$L.mPFC` | `r searchSpace$L.PCC` |
| Right       | `r searchSpace$R.mPFC` | `r searchSpace$R.PCC` |

>>>>

# In case I want to visualize stuff in the brain
labelCoords_vertex <- read.csv2('labelCoords_vertex.csv', sep = ",")[, 2:6]
labelCoords_vertex <- transform(labelCoords_vertex, x = as.numeric(as.character(x)), y = as.numeric(as.character(y)), z = as.numeric(as.character(z)))

@@

Figure 2 shows a representative partitioning of the search space for a single participant (100307; additional examples are presented on Supplemental Figure 1). The SP algorithm subdivides a network according to the positive versus negative values in the Fiedler vector (the eigenvector related to the second-to-lowest eigenvalue of the network's Laplacian matrix, see Methods). Since this is a data-driven approach, there is no a priori labeling for the two communities. We thus defined the DMN community as the one containing the one DMN-specific PCC parcel from the meta-analysis (7m). In qualitative terms, the resulting patterns contained substantial DMN coverage in posterior PCC (as dictated by our labeling strategy), with non-DMN vertices in anterior PCC. The mPFC region tended to include DMN vertices in its ventral-anterior and dorsal-anterior areas, with a persistent non-DMN pocket between them. This non-DMN section extended posteriorly into pregenual cinglate cortex (area a24).

@@

# Median size of the extra (>2) communities estimated by modularity
modularityCounts <- lapply(Summaries, function(data) {sort(plyr::count(as.character(data$Membership))$freq, decreasing = T)})
modularityCounts_two <- sapply(modularityCounts, function(x) {sum(x[seq(2)], na.rm = T)})
modularityCounts_extra <- sapply(modularityCounts, function(x) {sum(x[3:4], na.rm = T)})

@@

This comparison showed that both clustering methods were in high agreement across subjects (mean ARI = `r round(mean(RI_all_within$Index), digits = 2)`, SD = `r round(sd(RI_all_within$Index), digits = 2)`). Modularity showed a tendency to produce additional communities (median = `r median(nModularity)`, range = `r range(nModularity)`). However, these communities encompassed a small number of vertices (median = `r median(modularityCounts_extra)`, IQR = `r quantile(modularityCounts_extra, 0.25)` - `r quantile(modularityCounts_extra, 0.75)`) compared to the principal two (median = `r median(modularityCounts_two)`, IQR = `r quantile(modularityCounts_two, 0.25)` - `r quantile(modularityCounts_two, 0.75)`, suggesting that the overall network is composed of two main communities. In addition, the estimation of the optimal number of cuts based on the rank of the first non-zero eigenvalue (see "Partition evaluation" for details) showed that `r prcntDissection` percent of the individual networks were best partitioned into two communities.

@@

### *Pattern variability over time*

@@

There is precedent in thhe literature for the idea that the magnitude (and not just the sign) of the Fiedler vector values conveys important information about the role of each node in the network [@Gkantsidis2003; @Tian2018]. Therefore, we asked whether the magnitude of these eigenvector values was associated with the stability of nodes in time. Specifically, we estimated the proportion of DMN affiliations per node as a function of Fiedler vector values, using a logistic mixed effects model (Figure 4). The model yielded a positive significant relationship between these features ($\beta$ = `r round(SC_FVvPDMN_modelSummary$coefficients[2,1], digits = 2)`, SE = `r round(SC_FVvPDMN_modelSummary$coefficients[2,2], digits = 2)`, *p* < 0.001), meaning that vertices with higher absolute Fiedler vector values were more persistent in their relationship with their corresponding community over time. We refer to this proportion of affiliations over time as the stability of a given node.

@@

Once again we found low alignment across individuals for PCC (mean = `r round(mean(Halves_PCCComparison[[2]]), digits = 2)`, SD = `r round(sd(Halves_PCCComparison[[2]]), digits = 2)`) and mPFC (mean = `r round(mean(Halves_PFCComparison[[2]]), digits = 2)`, SD = `r round(sd(Halves_PFCComparison[[2]]), digits = 2)`), but both areas showed comparatively high levels of within-individual agreement (PCC: mean = `r round(mean(Halves_PCCComparison[[1]]), digits = 2)`, SD = `r round(sd(Halves_PCCComparison[[1]]), digits = 2)`; mPFC: mean = `r round(mean(Halves_PFCComparison[[1]]), digits = 2)`, SD = `r round(sd(Halves_PFCComparison[[1]]), digits = 2)`). We calculated an index of relative specificity by computing the ratio of each individual’s across-day (within-participant) ARI to the mean of all between-participant ARI values involving that individual. The index is expected to take on a value near 1 if partitionings are well aligned across individuals and/or are subject to a common level of measurement noise. It is expected to exceed 1 insofar as functional network organization is reliable and individual-specific. This index is intended to factor out the potential effects of measurement noise, as they are asumed to be constant for each region.

Figure 6 shows these ARI ratios for PCC and mPFC for all individuals As expected, we found that while both regions showed evidence for specificity (i.e. ratios > 1), the ratios for mPFC (median = `r round(median(Halves_RIRatios_PFC), digits = 2)`; IQR = `r quantile(Halves_RIRatios_PFC, 0.25)` - `r quantile(Halves_RIRatios_PFC, 0.75)`) were significantly higher than those for PCC (median = `r round(median(Halves_RIRatios_PCC), digits = 2)`; IQR = `r quantile(Halves_RIRatios_PCC, 0.25)` - `r quantile(Halves_RIRatios_PCC, 0.75)`) when compared with a paired permutation test (*p* `r ifelse(Halves_mPFCvPCC_permTest$Pval < 0.001, "< 0.001", paste("=", Halves_mPFCvPCC_permTest$Pval))`; Cohen's D = `r round(Halves_mPFCvPCC_ES, digits = 2)`) **(TBD: test against 1. I think it might be worth doing it for PCC's sake, as mPFC's numbers + figure already convey its idiosyncrasy very well. If we do it here, we'll do it for the run-specific analysis too)**. These test/retest results suggest that the topographic variability seen in mPFC arises at least in part from stable and subject-specific organizational patterns.

@@

The computational demands of our approach escalate quickly as the number of vertices increases. We thus explored the possible advantage of community detection relative to a simple seed-based functional connectivity analysis for estimating the individual-specific functional topography of mPFC. We examined whether maps generated with SP were more similar per participant across days than those computed from seed-based correlations. We generated a seed time-series by averaging all vertices in the PCC region of our search space, and calculated its correlation with the activity of each vertex in mPFC limiting this to 7m as a seed. I think it makes sense, especially since we're saying that PCC (1) is also subdivided to some extent, and (2) also shifts across individuals to some extent. Alternatively, you wouldn't use 7m as a seed unless you just wanted to find DMN. Full PCC is what a traditional analysis might have done. We compared the map of correlation values in mPFC to the map of unbinarized Fiedler vector values. We performed pairwise spatial Spearman correlations among maps computed for each day and method from all subjects. For `r nrow(unCorrParticipants %>% filter(Corrs < 0))` subjects, the maps produced with one of the days' datasets insufficiently covered area 7m, and so the community labeling was reversed (for one additional subject, both daily maps were reverted). ARI is robust to these labeling issues, but this produced strong negative correlations of the Fiedler vector for these subjects, even though the topography was highly similar. We thus visually inspected these subjects (n = 7) and reverted the labeling when it was clear that the topologies aligned across days. Figure 7 (left) shows that these pairwise comparisons resemble those from the day-to-day comparisons above, and suggests good alignment between methods, but particularly high agreement within subject and method.

@@

The thresholded partitions in our work qualitatively resembled the topographical arrangement of DMN sub-networks A and B proposed by Braga and Buckner [-@Braga2017]. We explored this similarity by reproducing their seed-based connectivity approach in three of our subjects. In their work, Braga and Buckner selected individual vertices along areas including the dorsolateral prefrontal cortex (DLPFC), temporo-parietal junction (TPJ), and parahippocampal cortex (PHC) that produced two anticorrelated networks [@Braga2017; @Braga2019]. Therefore, we hypothesized that if the communities we identified on the medial wall matched their proposed networks, we should be able to reproduce our prtitionings by selecting seeds in one these diagnostic areas. We confirmed seed placement based on the pattern observed in TPJ and PHC. The whole-brain seed-based functional connectivity maps for these subjects are juxtaposed to their spectral partitioning counterparts on figure 8.

Visual inspection of these networks shows a remarkable similarity between our DMN and non-DMN communities and the previously reported sub-networks A and B. Since our search space did not contain areas TPJ and PHC, lack of signal from these areas could drive the small discrepancies in the topography between our work and theirs (particularly for network A). Nevertheless, these qualitative similarities are important, as they collectively suggest there may exist a true functional partitioning of canonical DMN regions into two distinct communities. Results from the present investigation suggest that the topographic organization of the communities is variable across a large sample of participants, especially in mPFC, and that the SP algorithm is a useful and reliable tool for identifying the two communities in individuals.

@@

### *Table 1. Parcels from Glasser et al. (2016) contained in the search space.*
| Hemisphere      |             mPFC             |              PCC            |
| :-------------- | :--------------------------- | :-------------------------- |
| Left            | `r sort(searchSpace$L.mPFC)` | `r sort(searchSpace$L.PCC)` |
| --------------- | ---------------------------- | --------------------------- |
| Right           | `r sort(searchSpace$R.mPFC)` | `r sort(searchSpace$R.PCC)` |
|                 |                              |                             |
